Methods to prioritise pop-up active transport infrastructure

### 2020-05-09 article.Rmd
Add figure 3

```{r load}
library(sf)
library(tidyverse)
library(tmap)
u = "https://github.com/cyipt/tempCycleways/releases/download/0.1/rsf_leeds.Rds"
r_original = readRDS(url(u))
ur = "https://github.com/cyipt/tempCycleways/releases/download/0.1/rtid.csv"
rtid = readr::read_csv(ur)
# uh = "http://media.nhschoices.nhs.uk/data/foi/Hospital.csv"
# download.file(uh, "uh.csv")
h = readr::read_delim("uh-edited.csv", delim = "|")
# nrow(h) # 1221 hospitals
h_clean = h %>% 
  mutate(Longitude = as.numeric(Longitude)) %>% 
  mutate(Latitude = as.numeric(Latitude)) %>% 
  filter(!is.na(Longitude)) %>% 
  filter(!is.na(Latitude))
hsf = h_clean %>% st_as_sf(coords = c("Longitude", "Latitude"), crs = 4326)
```

```{r parameters}
min_pctgov = 0
min_grouped_pctgov = 100
min_grouped_length = 500
city_centre_buffer_radis = 8000
city_centre = tmaptools::geocode_OSM("leeds", as.sf = TRUE)
city_centre_buffer = stplanr::geo_buffer(city_centre, dist = 8000)
h_leeds = hsf[city_centre_buffer, ]
```

```{r preprocess}
# remove motorways
r_original$highway2 = r_original$highway
r_original$highway2 = gsub(pattern = "_link", replacement = "", r_original$highway)
highway_table = table(r_original$highway)
highway_rare = highway_table[highway_table < nrow(r_original) / 100]
highway_remove = names(highway_rare)[!grepl(pattern = "motor|living|ped", x = names(highway_rare))]
r_cleaned = r_original %>% 
  filter(!grepl(pattern = "motorway", x = highway)) %>% 
  mutate(highway2 = case_when(
  highway2 %in% highway_remove ~ "other",
  grepl(pattern = "path|track|trunk", highway2) ~ "other",
  grepl(pattern = "ped|liv", highway2) ~ "pedestrian/living_street",
  TRUE ~ highway2
  )) %>% 
  mutate(maxspeed = case_when(
    maxspeed <= 20 ~ "20 mph or less",
    maxspeed > 20 & maxspeed <= 30 ~ "30 mph",
    maxspeed > 40 & maxspeed <= 40 ~ "40 mph",
    maxspeed > 50 ~ "50+ mph",
  )) %>% 
  mutate(pctgov = as.numeric(pctgov))
rj = inner_join(r_cleaned, rtid)
cy = r_cleaned %>% filter(highway == "cycleway")
r_central = rj[city_centre_buffer, ]
r_main = r_central %>% 
  filter(grepl(pattern = "cycleway|primary|second|tert", highway2))
r_main_region = rj %>% 
  filter(grepl(pattern = "cycleway|primary|second|tert", highway2))
```

@@

```{r t1, results='asis'}
# rj %>%
#   st_drop_geometry() %>%
#   select(name, highway2, maxspeed, pctgov, width) %>%
#   table1::table1(~ highway2 + pctgov + width | maxspeed, data = ., )
knitr::include_graphics("https://user-images.githubusercontent.com/1825120/81471440-0aa1ba00-91e9-11ea-8e6e-dd9f7669bf98.png")
```

@@

@@

The region of analysis may seem like a basic consideration and for some cities it is.
However, there are various reasons why simply analysing and plotting all possible transport network segments within the city or regional boundaries may not be a good idea, as shown in Figure \@ref(fig:geographic-subsetting).

@@

```{r levels}
r_pct_lanes = r %>% 
  filter(pctgov > min_pctgov) %>% 
  filter((lanespsvforward + lanesforward) > 1 |
      (lanespsvbackward + lanesbackward) > 1
  )
touching_list = st_touches(r_pct_lanes)
g = igraph::graph.adjlist(touching_list)
components = igraph::components(g)
r_pct_lanes$group = components$membership
r_pct_lanes = r_pct_lanes %>% 
  group_by(group) %>% 
  mutate(group_length = sum(length)) %>% 
  mutate(pctgov_mean = mean(pctgov, na.rm = TRUE)) %>% 
  filter(pctgov_mean > min_grouped_pctgov)
r_pct_lanes = r_pct_lanes %>% filter(group_length > min_grouped_length)
r_filter_before_grouping = r %>% 
  filter(pctgov > min_pctgov) %>% 
  filter((lanespsvforward + lanesforward) > 1 |
           (lanespsvbackward + lanesbackward) > 1
  ) %>% 
  filter(pctgov > min_grouped_pctgov) %>% 
  filter(length > 100)
tmap_mode("plot")
m1 = tm_shape(city_centre_buffer) + tm_borders(col = "grey") +
  tm_shape(r_filter_before_grouping) + tm_lines() +
  tm_layout(title = "Filter then group (length > 100, cycling_potential > 100)")
m2 = tm_shape(city_centre_buffer) + tm_borders(col = "grey") +
  tm_shape(r_pct_lanes) + tm_lines() +
  tm_layout(title = "Group then filter (length > 500, cycling_potential > 100)")
ma = tmap_arrange(m1, m2)
ma
```

>>>>

```{r load}
library(sf)
library(tidyverse)
library(tmap)
u = "https://github.com/cyipt/tempCycleways/releases/download/0.1/rsf_leeds.Rds"
r_original = readRDS(url(u))
ur = "https://github.com/cyipt/tempCycleways/releases/download/0.1/rtid.csv"
rtid = readr::read_csv(ur)
# uh = "http://media.nhschoices.nhs.uk/data/foi/Hospital.csv"
# download.file(uh, "uh.csv")
h = readr::read_delim("uh-edited.csv", delim = "|")
# nrow(h) # 1221 hospitals
h_clean = h %>% 
  mutate(Longitude = as.numeric(Longitude)) %>% 
  mutate(Latitude = as.numeric(Latitude)) %>% 
  filter(!is.na(Longitude)) %>% 
  filter(!is.na(Latitude))
hsf = h_clean %>% st_as_sf(coords = c("Longitude", "Latitude"), crs = 4326)
region = ukboundaries::leeds
```

```{r parameters}
min_pctgov = 0
min_grouped_pctgov = 100
min_grouped_length = 500
city_centre_buffer_radius = 8000
city_centre = tmaptools::geocode_OSM("leeds", as.sf = TRUE)
city_centre_buffer = stplanr::geo_buffer(city_centre, dist = city_centre_buffer_radius)
h_leeds = hsf[city_centre_buffer, ]
```

```{r preprocess}
# remove motorways
r_original$highway2 = r_original$highway
r_original$highway2 = gsub(pattern = "_link", replacement = "", r_original$highway)
highway_table = table(r_original$highway)
highway_rare = highway_table[highway_table < nrow(r_original) / 100]
highway_remove = names(highway_rare)[!grepl(pattern = "motor|living|ped", x = names(highway_rare))]
r_cleaned = r_original %>% 
  filter(!grepl(pattern = "motorway", x = highway)) %>% 
  mutate(highway2 = case_when(
  highway2 %in% highway_remove ~ "other",
  grepl(pattern = "path|track|trunk", highway2) ~ "other",
  grepl(pattern = "ped|liv", highway2) ~ "pedestrian/living_street",
  TRUE ~ highway2
  )) %>% 
  mutate(maxspeed = case_when(
    maxspeed <= 20 ~ "20 mph or less",
    maxspeed > 20 & maxspeed <= 30 ~ "30 mph",
    maxspeed > 30 ~ "40+ mph",
  )) %>% 
  mutate(pctgov = as.numeric(pctgov)) 
table(r_cleaned$maxspeed)
rj = inner_join(r_cleaned, rtid) %>% 
  mutate(lanes_f = lanespsvforward + lanesforward) %>% 
  mutate(lanes_b = lanespsvbackward + lanesbackward) %>% 
  mutate(`N. lanes` = lanes_f + lanes_b)
cy = r_cleaned %>% filter(highway == "cycleway")
r_central = rj[city_centre_buffer, ]
r_main = r_central %>% 
  filter(grepl(pattern = "cycleway|primary|second|tert", highway2))
r_main_region = rj %>% 
  filter(grepl(pattern = "cycleway|primary|second|tert", highway2))
```

@@

```{r t1, results='asis'}
# rj %>%
#   st_drop_geometry() %>%
#   select(name, highway2, maxspeed, pctgov, width) %>%
#   table1::table1(~ highway2 + pctgov + width | maxspeed, data = ., )
knitr::include_graphics("https://user-images.githubusercontent.com/1825120/81476361-8c094480-9209-11ea-97d4-f4ad09e96d69.png")
```

@@

<!-- Could say more about the case study city here if there is space -->

@@

The region of analysis may seem like a basic consideration: most cities have well-defined administrative zones.
In Leeds and many other cities, it makes sense to focus on the region directly surrounding the city centre, in a kind of 'geographical triage' to omit from the analysis pop-up options in the outskirts and focus valuable attention on the routes that are most likely to serve the highest number of people.
We set a modifiable parameter `city_centre_buffer_radius` to 8km (5 miles) as a 'first pass' for the case study city of Leeds.
The results are shown 
However, there are various reasons why simply analysing and plotting all possible transport network segments within the city or regional boundaries may not be a good idea, as shown in Figure \@ref(fig:gsub).

```{r gsub, fig.cap="Illustration of geographic subsetting based on distance to the city centre rather than based on location within somewhat arbitrarily shaped city boundaries"}
# nrow(r_central)
# nrow(rj)
square = stplanr::geo_bb(shp = city_centre_buffer)
m1 = tm_shape(r_main_region) +
  tm_lines("maxspeed", palette = "-Set1", lwd = "N. lanes", scale = 5,
           lwd.legend = c(1, 2, 3, 4, 5, 6)) +
  tm_shape(region) + tm_borders() +
  tm_shape(square) + tm_borders(lty = 3) +
  tm_scale_bar() 
# m1
m2 = tm_shape(r_main) +
  tm_lines("maxspeed", palette = "-Set1", lwd = "N. lanes", scale = 5,
           lwd.legend = c(1, 2, 3, 4, 5, 6)) +
  tm_shape(city_centre_buffer) + tm_borders() +
  tm_scale_bar() +
  tm_layout(legend.show = FALSE)
tmap_arrange(m1, m2, nrow = 1)
```

@@

```{r levels}
r_pct_lanes = rj %>% 
  filter(pctgov > min_pctgov) %>% 
  filter((lanespsvforward + lanesforward) > 1 |
      (lanespsvbackward + lanesbackward) > 1
  )
touching_list = st_touches(r_pct_lanes)
g = igraph::graph.adjlist(touching_list)
components = igraph::components(g)
r_pct_lanes$group = components$membership
r_pct_lanes = r_pct_lanes %>% 
  group_by(group) %>% 
  mutate(group_length = sum(length)) %>% 
  mutate(pctgov_mean = mean(pctgov, na.rm = TRUE)) %>% 
  filter(pctgov_mean > min_grouped_pctgov)
r_pct_lanes = r_pct_lanes %>% filter(group_length > min_grouped_length)
r_filter_before_grouping = rj %>% 
  filter(pctgov > min_pctgov) %>% 
  filter((lanespsvforward + lanesforward) > 1 |
           (lanespsvbackward + lanesbackward) > 1
  ) %>% 
  filter(pctgov > min_grouped_pctgov) %>% 
  filter(length > 100)
tmap_mode("plot")
m1 = tm_shape(city_centre_buffer) + tm_borders(col = "grey") +
  tm_shape(r_filter_before_grouping) + tm_lines() +
  tm_layout(title = "Filter then group (length > 100, cycling_potential > 100)")
m2 = tm_shape(city_centre_buffer) + tm_borders(col = "grey") +
  tm_shape(r_pct_lanes) + tm_lines() +
  tm_layout(title = "Group then filter (length > 500, cycling_potential > 100)")
ma = tmap_arrange(m1, m2)
ma
```
